---
title: "shiny_workshop"
author: "Sebastien Renaut"
date: '2018-09-12'
output: html_document
runtime: shiny
---

```{r setup, echo=T}
#install.packages("shiny")
library(shiny)
```

# What is Shiny?
• This R Markdown document is made interactive using `shiny`. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

• To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs
• You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.   

• The code below demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to generate plots.  


```{r eruptions, echo=T}
#Here we will use the faithful geyser example dataset to generate histogram and a density function.
#Go see what the data looks like
head(faithful)

inputPanel(
  #selecting the type of input (n_breaks)
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  #selecting the type of input (bw_adjust)
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

  #What plot will you make
renderPlot({
  #an histogram of faithful$eruptions with breaks defined according to above
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  #a density line of faithful$eruptions with adjust defined according to above
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application
• It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function.  

```{r tabsets, echo=T}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

• Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

• In all of R code chunks above the `echo = T` attribute is used. This is to show the R code (but obviously should be set to `echo = F` if used on a website).


## Another (simple) example
• See [gallery](https://shiny.rstudio.com/gallery/) for more examples.

```{r mtcars, echo=T} 
# Define UI for miles per gallon app ----
ui <- fluidPage(

  # App title ----
  titlePanel("Miles Per Gallon"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Selector for variable to plot against mpg ----
      selectInput("variable", "Variable:",
                  c("Cylinders" = "cyl",
                    "Transmission" = "am",
                    "Gears" = "gear"))
    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Formatted text for caption ----
      h3(textOutput("caption")),

      # Output: Plot of the requested variable against mpg ----
      plotOutput("mpgPlot")

    )
  )
)

# Define server logic to plot various variables against mpg ----
server <- function(input, output) {

  # Compute the formula text ----
  # This is in a reactive expression since it is shared by the
  # output$caption and output$mpgPlot functions
  formulaText <- reactive({
    paste("mpg ~", input$variable)
  })

  # Return the formula text for printing as a caption ----
  output$caption <- renderText({
    formulaText()
  })

  # Generate a plot of the requested variable against mpg ----
  # and only exclude outliers if requested
  output$mpgPlot <- renderPlot({
    plot(as.formula(formulaText()),
            data = mtcars,
            col = "#75AADB", pch = 19)
  })

}

# Return a Shiny app object
shinyApp(ui = ui, server = server,options=list(height = 550))
```  

# Deployment
• Shiny documents must be supported by a live R session running behind the scenes. When you run a Shiny document locally, it uses your local R session. Only you can see and interact with the document. 

• If you want to share the document with people who do not have R installed, or do not want to run the document locally, you will have to deploy the document on a server, and share the URL of the document. Then other people only need a web browser to visit your document.

• Two ways to deploy a Shiny document: You can either use a hosted service provided by **RStudio**, or set up your own server. The first way is technically easier, but sometimes you may not be allowed to use an external hosted service, so you have to install the required software (Shiny Server or RStudio Connect) on your own server to deploy the Shiny documents.

# ShinyApps.io
• Sign up to http://www.shinyapps.io/ 

• Connect to the remote server

`#install.packages("rsconnect")`  
`library(rsconnect)`  
`rsconnect::setAccountInfo(name='sebastien',token='this_is_my_token_please_setup_your_own',secret='this_is_my_secret_please_setup_your_own')` 

• Publish document at the top right corner  


