---
title: "shiny_workshop"
author: "Sebastien Renaut"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    number_sections: T
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: T
runtime: shiny
---

```{r setup, echo=T}
#install.packages("shiny")
library(shiny)
```

# What is Shiny?
* This `R Markdown'`html document is made interactive using `shiny`. 

* To reate documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

* To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

# Set-up

## Header
>\---  
>title: "shiny_workshop"  
>author: "Sebastien Renaut"  
>date: "`r format(Sys.time(), '%B %d, %Y')`"  
>output: html_document  
>runtime: shiny  
>\---  

## View document with shiny  
* Run document (instead of Knit)

## Inputs and Outputs
* Outputs are automatically updated whenever inputs change.   

* The code below demonstrates how a standard plot can be made interactive by wrapping it in the Shiny `renderPlot` function (*output*). The `selectInput` and `sliderInput` functions create the *input* widgets used to generate plots.  


```{r eruptions, echo=T}
#Here we will use the faithful geyser example dataset to generate histogram and a density function.
#Go see what the data looks like
head(faithful)

#What input will you give to your function:
inputPanel(
  #selecting the type of input (n_breaks)
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  #selecting the type of input (bw_adjust)
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

  #What plot will you make:
renderPlot({
  #an histogram of faithful$eruptions with breaks defined according to above
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  #a density line of faithful$eruptions with adjust defined according to above
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application
* It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function.  

```{r tabsets, echo=T}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

* Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.  

## Another (simple) example
* See [gallery](https://shiny.rstudio.com/gallery/) for more examples.

```{r mtcars, echo=T} 
# Define UI for miles per gallon app ----
ui <- fluidPage(

  # App title ----
  titlePanel("Miles Per Gallon"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Selector for variable to plot against mpg ----
      selectInput("variable", "Variable:",
                  c("Cylinders" = "cyl",
                    "Transmission" = "am",
                    "Gears" = "gear"))
    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Formatted text for caption ----
      h3(textOutput("caption")),

      # Output: Plot of the requested variable against mpg ----
      plotOutput("mpgPlot")

    )
  )
)

# Define server logic to plot various variables against mpg ----
server <- function(input, output) {

  # Compute the formula text ----
  # This is in a reactive expression since it is shared by the
  # output$caption and output$mpgPlot functions
  formulaText <- reactive({
    paste("mpg ~", input$variable)
  })

  # Return the formula text for printing as a caption ----
  output$caption <- renderText({
    formulaText()
  })

  # Generate a plot of the requested variable against mpg ----
  # and only exclude outliers if requested
  output$mpgPlot <- renderPlot({
    plot(as.formula(formulaText()),
            data = mtcars,
            col = "#75AADB", pch = 19)
  })

}

# Return a Shiny app object
shinyApp(ui = ui, server = server,options=list(height = 550))
```  

# Deployment
* When you run a `shiny` document locally, it uses your local R session. Only you can see and interact with the document. 

* The power of `shiny` comes from a live R session running behind the scenes to all allow users to interct with documents.   

* You will have to deploy the document on a server, and share the URL of the document. Then other people only need a web browser to visit your document.

* Two ways to deploy a `shiny` document: 
  + Host it by service provided by **RStudio**
    + Technically easier.
    + webpage hosted on shinyapps.io
    + Freemium businness model
  + Set up your own server
    +  Independence from *Rstudio*

# ShinyApps.io
* Sign up to http://www.shinyapps.io/ 

* Connect to the remote server (may need to install `rsconnect` and `packrat`)

*Also need to set remote account info
  + Once signed up, go to http://www.shinyapps.io/ >Account >Tokens >Show   

>  
>rsconnect::setAccountInfo(name='yourname',token='alphanumeric',secret='ASCII') 
>

* Publish document in Rtudio by clicking on *blue eye* at the top right corner. Voilà!  


